Uncovering underlying high-level musical content in the temporal domain \\ \small Leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn music embeddings with applications to MIR downstream tasks

### 2023-05-30 discussion/discussion.tex
Updates from Overleaf

In this work, we introduced a method to uncover underlying high-level musical
content in the temporal domain by leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn music latent representations with applications to MIR downstream tasks. Building on top of existing approaches and architectures and wrapped within MSAF \cite{NietoMSAF:FRAMEWORK}, we replaced traditional features with deep embeddings trained to represent high-level musical content information.

>>>>

In this work, we introduced a method to uncover underlying high-level musical
content in the temporal domain by leveraging self-supervised deep neural networks, inductive bias, and aural skills to learn music latent representations with applications to MIR downstream tasks. Building on existing approaches and architectures, we replaced traditional features with deep embeddings trained to represent high-level musical content information.

### 2023-05-30 evaluation/evaluation.tex
Updates from Overleaf

\input{evaluation/subsections/beyond dsp}

>>>>

\section{Music Structure Analysis as downstream task}

While the usefulness of deep audio embeddings can be evaluated in countless downstream tasks, we have chosen Music Structure Analysis (MSA) for its popularity, complexity, and product-compelling nature. This interdisciplinary field aims to understand the structure of music \cite{Nieto2020Audio-BasedApplications}. Due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces challenges like boundary placement ambiguity and similarity quantification \cite{NietoPerceptualMusic}.

All the evaluation experiments have been wrapped around MSAF \cite{NietoMSAF:FRAMEWORK}, a popular framework for Music Structure Analysis.

All the evaluation metrics have been computed using the mir-eval package \cite{RaffelMir_eval:METRICS}.

\subsection{SALAMI dataset}

SALAMI (Structural Analysis of Large Amounts of Music Information) aims to conduct massive structural analyses of different types of music \cite{Smith2011DESIGNANNOTATIONS}. The project covers diverse music, from Western pop to Indian classical and from live to studio recordings.

The authors propose a unique music analysis approach that segments pieces into sections rather than traditional forms. Their method combines perceptual, functional, and transcription analyses, considers music's hierarchical nature, and uses markers to determine analysis timescales. While providing a comprehensive investigation, each analysis approach has limitations in accurately reflecting musical structure, labeling identical or different music sections, and capturing the complete structural framework. This innovative approach aims for a more nuanced, holistic music analysis.

\section{\textit{Embeddiogram} feature}

The \textit{embeddiogram} is derived by applying our pre-trained neural network model to sliding windowed segments across the audio signal, generating a sequence of embedding vectors. These relatively high-dimensional embedding vectors collectively form a two-dimensional description of the audio signal's musical content.

Below is a detailed explanation of how we compute the \textit{embeddiogram} from a given audio signal. This process comprises five key steps, including loading the audio data, slicing the audio data into windowed segments, processing each window using a pre-trained model to produce an embedding, collecting these embeddings, and finally normalizing the Embeddiogram.

\begin{enumerate}
\item \textbf{Load the audio data}: The audio data is loaded into memory as a one-dimensional array of length $N$.

\item \textbf{Slice the audio data}: The audio data is segmented into overlapping windows. Each window contains $w$ samples, and consecutive windows are separated by $h$. This gives a total of $H$ windows, defined as:
\begin{equation}
H = 1 + \left\lfloor \frac{N - w}{h} \right\rfloor
\end{equation}
In the case where $\left( N - w \right) \mod h > 0$, we have $H += 1$ to account for the final, potentially smaller, window.

\item \textbf{Process each window}: Each window of audio data is processed independently, passed through the pre-trained neural network model and transformed into an embedding vector. Formally, for each window $w_i$ of audio data, we have:
\begin{equation}
\text{embedding}_i = \text{model}(w_i)
\end{equation}

\item \textbf{Collect the embeddings}: The embedding vectors are collected and stacked together. Each row represents a feature vector for a given time frame to form the \textit{embeddiogram}, denoted as $\text{e}$:
\begin{equation}
\text{e} = \begin{bmatrix} \text{embedding}_1 \\ \text{embedding}_2 \\ \vdots \\ \text{embedding}_H \end{bmatrix}
\end{equation}

\item \textbf{Normalize the \textit{embeddiogram}}: The \textit{embeddiogram} is normalized to have a minimum value of 0 and a maximum value of 1. The normalization process is given by:
\begin{equation}
E'_{ij} = \frac{e_{ij} - \min(E)}{\max(E) - \min(E)}
\end{equation}
\end{enumerate}

\begin{figure}[ht]
  \centering
  \begin{minipage}[b]{.8\linewidth}
    \centering
    \includegraphics[width=\linewidth]{figures/images/salami_391_embeddiogram.png}
    \caption[]{Caption for image1}
    \label{fig:image1}
  \end{minipage}

  \begin{minipage}[b]{.8\linewidth}
    \centering
    \includegraphics[width=\linewidth]{figures/images/samali_391_novelty_curve_peaks.png}
    \caption[]{Caption for image2}
    \label{fig:image2}
  \end{minipage}
\end{figure}

\begin{figure}[ht]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/salami_391_R matrix after Gaussian smoothing.png} % first figure itself
        \caption[]{first figure}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/salami_391_Lag matrix after Gaussian smoothing.png} % second figure itself
        \caption[]{second figure}
    \end{minipage}
    \vfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/salami_391_Q_matrix.png} % third figure itself
        \caption[]{third figure}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/salami_391_binary transitive matrix.png} % fourth figure itself
        \caption[]{fourth figure}
    \end{minipage}
\end{figure}

### 2023-05-30 introduction/Subsections/Assumptions.tex
Updates from Overleaf

This thesis posits the existence of high-level musical concepts invariant to sonic qualities that evolve and unfolds through time, such as melody, harmony, rhythm, form, motif, counterpoint, etc. This idea parallels the nature of the symbolic domain (sheet music), which maintains its essence despite being interpreted through diverse performances using various instruments and styles. While a single waveform might have a limited number of sheet music representations that adequately capture its musical components, a single piece can be interpreted in infinite ways through different instruments, voices, tempos, and interpretations.

@@

This concept finds grounding in the theories of Kurt Koffka, a founder of the Gestalt school of psychology, who proposed a holistic approach to understanding complex forms, contrary to the structuralist practice of reducing mental processes into essential elements \cite{Koffka2013PrinciplesPsychology}. When applied to music, the principles of Gestalt psychology offer insights into how we perceive, organize, and interpret musical patterns. They suggest that our minds perceive auditory input much like visual input, seeking patterns and structures. This understanding significantly contributes to the comprehension of cognitive processes in musical perception and organization, influencing music theory, cognition, and therapy \cite{Lerdahl1985AMusic}.

>>>>

This thesis proposes the existence of high-level musical concepts such as melody, harmony, rhythm, form, motif, and counterpoint. These concepts are invariant to sonic qualities, demonstrating their independence and endurance over time. Furthermore, we argue these elements interact and evolve with each other. This idea parallels the nature of the symbolic domain (sheet music), which maintains its essence despite being interpreted through diverse performances using various instruments and styles. While a single waveform might have a limited number of sheet music representations that adequately capture its musical components, a single piece can be interpreted in infinite ways through different instruments, voices, tempos, and interpretations.

@@

This concept finds grounding in the theories of Kurt Koffka, a founder of the Gestalt school of psychology, who proposed a holistic approach to understanding complex forms, contrary to the structuralist practice of reducing mental processes into essential elements \cite{Koffka2013PrinciplesPsychology}. When applied to music, the principles of Gestalt psychology suggest that our minds perceive auditory input much like visual input, seeking patterns and structures. This understanding significantly contributes to the comprehension of cognitive processes in musical perception and organization, influencing music theory, cognition, and therapy \cite{Lerdahl1985AMusic}.

### 2023-05-30 methods/subsections/implementation.tex
Updates from Overleaf

While we have experimented with helpful audio augmentation tools such as \cite{Spijkervet2021Spijkervet/torchaudio-augmentations:V1.0} and \cite{Kharitonov2020DataDomain}, the specific requirements of our experiments necessitated the development of our transformation chain using \textit{torchaudio} \cite{Yang2021TorchAudio:Processing}: given an anchor audio signal $A[n]$, we generate a positive signal $P[n]$ by applying a series of amplitude, time-domain, frequency-domain, modulation, reverberation, and nonlinear effects with additive noise on top of it.

>>>>

While we have experimented with helpful audio augmentation tools such as \cite{Spijkervet2021Spijkervet/torchaudio-augmentations:V1.0} and \cite{Kharitonov2020DataDomain}, the specific requirements of our experiments necessitated the development of our transformation chain using \textit{torchaudio's} \cite{Yang2021TorchAudio:Processing} implementation of \textit{SoX} \cite{sox}: given an anchor audio signal $A[n]$, we generate a positive signal $P[n]$ by applying a series of amplitude, time-domain, frequency-domain, modulation, reverberation, and nonlinear effects with additive noise on top of it.

### 2023-05-30 results/results.tex
Updates from Overleaf

\input{figures/neural networks/manifold}

@@

\begin{figure}[!ht]
\includegraphics[clip,width=\columnwidth]{figures/PlotTimeSeriesResult}% 
\caption[Graph.]{This is an example of a figure and its caption.}
\label{fig:timeseries}
\end{figure}

>>>>

@@

Extend table by Hernandez-Olivan, C., Beltran, J. R. \& Diaz-Guerra, D. 
\cite{Hernandez-Olivan2021MusicFeatures}

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{5cm}XXXXX}
\toprule
\thead{\textbf{Year, Authors, Ref}} & \thead{\textbf{Algorithm}} & \thead{\textbf{Input}} & \thead{\textbf{Method}} & \thead{\textbf{F-measure SALAMI}} \\
\midrule
\addlinespace
2009, Paulus \& Klapuri, [24] & PK & MFCCs, chromas & Fitness function  & - \\
\addlinespace
2010, Mauch et al., [25] & MND1 & MFCCs, Discrete Cepstrum & HMM  & - \\
\addlinespace
2011, Sargent et al., [26]& SB-VRS1 & Chords estimation & Viterbi  & - \\
\addlinespace
2012, Kaiser et al., [27]& KSP2 & SSM & Novelty measure  & 0.286 \\
\addlinespace
2013, McFee \& Ellis, [20] & MP2 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
2014, Nieto \& Bello, [28] & NB1 & MFCCs + chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
2015, Cannam et al., [29] & CC1 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
2016, Nieto, [30] & ON2 & Constant-Q Transform Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
2017, Cannam et al., [29] & CC1 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
2007, Turnbull et al., [19] & - & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
2011, Sargent et al., [34] & - & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
2014, Ullrich et. al, [22] & - & MLS & CNN  & 0.465 \\
\addlinespace
2015, Grill \& Schlüter, [4] & - & MLS + SSLMs & CNN  & 0.523 \\
\addlinespace
2015, Grill \& Schlüter, [5] & - & MLS + PCPs + SSLMs & CNN  & 0.508 \\
\addlinespace
2017, Hadria \& Peeters, [35] & - & MLS + SSLMs & CNN  & 0.291 \\
\bottomrule
\end{tabularx}
\caption{Your Caption}
\label{tab:my_label}
\end{table}
